Polytope notes

### 2020-02-21 notes/bo-separated.tex

\begin{theorem}
  Let $L(\cdot;p)$ be a polyhedral function with $U^* := \arg\min_{u} L(u;p)$.
  Then there exists $c > 0$ such that $L(u;p) \geq c \cdot \|U^* - u\|_2$ for all $u$.
\end{theorem}
\begin{proof}
  Consider the epigraph of $L(\cdot;p)$, a polyhedral set (intersection of finitely many halfspaces).
  Recall that each face $F_S$ of the epigraph is uniquely identified by a subset of constraints $S$ which are tight for exactly that face.
  In particular, $U^*$ is a face (as it is the set of points exposed by the ``downward'' direction), identified by some subset $S^*$.

@@

>>>>

\begin{theorem}
  Let $L(\cdot;p)$ be a polyhedral function with $U^* := \arg\min_{u} L(u;p)$ and $L^* := \min_u L(u;p)$.
  Then there exists $c > 0$ such that $L(u;p) \geq L^* + c \cdot d(U^*,u)$ for all $u$.
\end{theorem}
\begin{proof}
  Consider the epigraph of $L(\cdot;p)$, a polyhedral set (intersection of finitely many halfspaces).
  Recall that each face $F_S$ of the epigraph is uniquely identified by a subset of constraints $S$ which are tight for exactly that face.
  In particular, $U^*$ is a face (as it is the set of points exposed by the ``downward'' direction), identified by some subset $S^*$.

@@

(Bo, 2020-02-11, attempt to rigorize previous proof sketch)

\begin{theorem}
  Let $L(\cdot;p)$ be a polyhedral function with $U^* := \arg\min_{u} L(u;p)$ and $L^* := \min_u L(u;p)$.
  Then there exists $c > 0$ such that $L(u;p) \geq L^* + c \cdot d(U^*,u)$ for all $u$.
\end{theorem}
\begin{proof}
  Let $U^*$ be the set of minimizers of $L(\cdot;p)$, a polyhedral set in $\reals^d$.
  Let $S^*$ be the set of constraints defining $U^*$, i.e. a minimal set of halfspaces whose intersection is $U^*$.
  For $S \subseteq S^*$, let $U_S$ be the face of $U^*$ for which the constraints $S$ are tight.
  Let $U_S'$ be the relative interior of the face $U_S$.

  Let $u \in U_S'$ and define the following sets.
  Let $V_S$ be the set of normal vectors exposing $u$.
  Let $D_S$ be the set of subgradients of $L(\cdot;p)$ at $u$.
  Assertion 1: These sets are the same for all $u \in U_S'$, justifying the notation.

  Assertion 2: the boundary of $U^*$ is the disjoint union of the relative interiors of the sets $U_S'$ for nonempty $S$.

  Now, for nonempty $S$, let $c_S := \inf_{v \in V_S} \sup_{d \in D_S} v \cdot d$.
  \textbf{Key claim:} $c_S > 0$.
  Proof: Write $f(v) := \sup_{d \in D_S} v \cdot d$.
  We claim $f$ is a continuous function of $v$ (Bo: TODO, justify!).
  Next, we claim $f(v) > 0$ for all $v \in V_S$.
  Otherwise, we would be able to obtain a contradiction for $u \in U_S'$ being on the boundary of the optimal set $U^*$ with normal vector $v$ and subgradient set $D_S$. (Bo: TODO actually obtain this contradiction.)

  Assertion 3: $V_S$ is a compact set.

  So $f$, a continuous function on a compact set, attains its infimum.
  Since $f(v) > 0$ for every $v \in V_S$, we have $c_S = \inf_{v \in V_S} f(v) > 0$.
  This proves the key claim.

  Now, we prove the theorem.
  Let $c = \min_{\emptyset \subsetneq S \subsetneq S^*} c_S$, noting that $c > 0$ as the minimum of a finite set of positive numbers.

  Fix any $u$.
  If $u \in U^*$, then the theorem's conclusion is trivially satisfied.
  So suppose $u \not\in U^*$.
  Let $u^*$ be the projection of $u$ onto $U^*$.
  We have $u^* \in U_S$ for some unique nonempty $S$ (by Assertion 2).
  We also have, for some $v \in V_S$, the identity $u = u^* + d(U^*,u) v$.
  Therefore, by definition of subgradients, we have
  \begin{align*}
    L(u;p) &\geq L^* + d(U^*,u) \sup_{d \in D_S} v \cdot d  \\
           &\geq L^* + d(U^*,u) c_S  \\
           &\geq L^* + d(U^*,u) c .
  \end{align*}
\end{proof}
